---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

```{r}
library(dplyr)
ac <- read_stats19_ac()
ca <- read_stats19_ca()
ve <- read_stats19_ve()

# merge all stats 19 data at the (lowest) vehicle level
all_stats19 <- dplyr::inner_join(ve, ca)
all_stats19 <- dplyr::inner_join(all_stats19, ac)
# now you can analyse the UK's stats19 data in a single table!
head(all_stats19$Casualty_Class)

# Just the cyclists - how many vehicles vs records?
acy <- all_stats19[all_stats19$Casualty_Type == "Cyclist",]
nrow(unique(acy[,1])) / nrow(acy)
summary(acy$Number_of_Vehicles)

acp <- all_stats19[all_stats19$Casualty_Type == "Pedestrian",]
nrow(unique(acp[,1])) / nrow(acp)
summary(acp$Number_of_Vehicles)
```

```{r}
# Joining to accidents first
as19 <- right_join(ac, ca)
as19 <- right_join(as19, ve)
```


## Which modes are responsible for casualties

```{r}
ca$AVR <- paste0(ca$Accident_Index, ca$Vehicle_Reference)
ve$AVR <- paste0(ve$Accident_Index, ve$Vehicle_Reference)
a$AVR <- paste0(a[1], a$Vehicle_Reference)
hitter <- !a$AVR  %in% ca$AVR
summary(hitter)
summary(a$Vehicle_Type[hitter])
```


```{r}
cav <- inner_join(ca, ac)
cav <- inner_join(cav, ve)
cav$Perpetrator_Casualty_Type <- as.factor(paste0(cav$Vehicle_Type, " hit a ", cav$Casualty_Type))
cavt <- table(cav$Perpetrator_Casualty_Type)
cavt <- cavt[order(cavt, decreasing = T)]
cavt[1:10]
sum(cav$Vehicle_Type == "Car" & cav$Casualty_Type == "Cyclist")
as199 <- as19[!is.na(as19$Vehicle_Type) & !is.na(as19$Casualty_Type),]
sum(as199$Vehicle_Type == "Car" & as199$Casualty_Type == "Cyclist")
```


## Who gets hit?

```{r}
names(ca)
(cat <- table(ca$Casualty_Type)[order(table(ca$Casualty_Type), decreasing = T)])
barplot(cat[1:3])

sel <- all_stats19$Casualty_Type == "Pedestrian" & all_stats19$Vehicle_Type == "Pedal cycle"
ss <- all_stats19[sel,]

sum(sel) / nrow(all_stats19)
sum(all_stats19$Vehicle_Type == "Car" & all_stats19$Casualty_Type == "Cyclist")

```





## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))