---
title: "Accident hotspots in Leeds"
author: "Robin Lovelace"
date: "May 29, 2015"
output: 
  pdf_document: 
    fig_caption: yes
---

```{r, include=FALSE}
pkgs <- c("grid", "png")
lapply(pkgs, library, character.only = T)
```

## Introduction

This represents a preliminary analysis of accident hotspots
for cyclists in Leeds.
It is preliminary because it does not include 2014 data and because
it does not normalise by estimated exposure. Also, it does not
make the distinction between different types of cycling trips
(primarily commuter trips, by time of day). This has the advantage of
including more data points. The disadvantage is that it heavily reflects
where cycling is greatest in Leeds, in the north of the city.

However, due to the linear nature of roads, we can clearly see that
some areas are more risky than others: it is fair to assume that cyclists
travel along the roads before and after the hotspots identified in this analysis.
The fact that they coincide with junctions and roundabouts, where we would
intuitely expect crashes to be more frequent, supports the analysis (Fig. 1).

```{r, echo=FALSE, fig.cap="Rasterised layer of accident hotspots with polygons representing clusters estimated using R."}
grid.raster(readPNG("../figures/leedshot-50m.png"))
```


All the analysis underlying this research is reproducible.
In addition, the data underlying the results can be accessed for free online
by following the code provided in the paper's repository: 
https://github.com/Robinlovelace/bikeR

In addition, key datasets have been placed on an interactive
web map so they can be explored.

## Bicycle crash hotspots

The analysis is based on STATS19 data collected during 9 years,
from the beginning of 2005 until the end of 2013. 
Overall there were 4,738 incidents reported by STATS19 in West Yorkshire
over this period, providing a sufficient amount of geocoded data to 
explore accident hotspots in the area. The script to load and process the
data can be found here: https://github.com/Robinlovelace/bikeR/blob/master/video-routes/load-stats19.R

The 16 worst offending hotspots, in terms of concentration of crashes
are displayed in Fig. 1 above and summarised in the table below.

Table: Worst accident hotspots in West Yorkshire arranged by accident
rate (incidents per hectare per year of hotspots with cell size of 50m).

| n_crashes|  area| accident_rate|name                                             |
|---------:|-----:|-------------:|:------------------------------------------------|
|         6|  1871|          3.57|Savins Mill Way, LS5                             |
|         4|  1354|          3.28|50 Headingley Lane, LS6 2BW                      |
|         9|  3556|          2.81|2 Burley Wood Mount, LS4 2QG                     |
|        10|  4146|          2.68|152 Otley Road, LS16 5LG                         |
|         7|  3740|          2.08|187 Otley Road, LS6 3QG                          |
|         9|  6084|          1.64|Bayswater Road, LS8 5NA                          |
|        10|  7021|          1.58|356 Kirkstall Road, LS4 2HQ                      |
|        11|  7795|          1.57|40 Queen's Road, LS6 1HU                         |
|        13|  9309|          1.56|2 Rawdon Road, Horsforth, LS18 5DZ               |
|        12|  8767|          1.52|6 Otley Road, LS16 6QR                           |
|         5|  3723|          1.49|293 Roundhay Road, LS8 4HS                       |
|         8|  6901|          1.29|University of Leeds, 197 Woodhouse Lane, LS2 9DY |
|        83| 87225|          1.06|2 Otley Road, LS6 2DG                            |
|        43| 46303|          1.03|40 Headingley Lane, LS6 2EB                      |
|        NA|    87|            NA|2 Bridge Road, LS5 3BL                           |
|        NA|  1478|            NA|105 Chapeltown Road, LS7 3SG    |

## Subsetting hotspots for further investigation

By observing Fig. 1 and Table 1, we identified 5 offending areas
to visit and film. The next stage is to see if study participants would find them
subjectively more dangerous than the surroundings. These are displayed in
Fig. 2 and in Table 2 below:

Table: 5 case study 'danger hotspots'

|name                         |reason                                         |
|:----------------------------|:----------------------------------------------|
|Savins Mill Way, LS5         |Has the highest accident rate                  |
|50 Headingley Lane, LS6 2BW  |High accident rate                             |
|356 Kirkstall Road, LS4 2HQ  |Moderate accident rate, isolated at a junction |
|2 Otley Road, LS6 2DG        |Largest area, surrounded by other hotspots     |
|105 Chapeltown Road, LS7 3SG |In an under-represented part of town, isolated |

```{r, echo=FALSE, fig.cap="5 hotspots of cycle crashes in north Leeds. See http://rpubs.com/RobinLovelace/83969 for interactive map."}
grid.raster(readPNG("../figures/5-hots-screenshot.png"))
```

## Further work

An issue with the analysis so far, as mentioned above, is that it takes no
account of exposure, i.e. estimates of the current rate of cycling.

This data can be estimated and allocated to the route network thanks
to recent developments such as the National Propensity to Cycle Tool
(NPCT). Normalising by this exposure would likely reveal hotspots in
the south of the city where  there is less cycling.

```{r, echo=FALSE, fig.cap="Cycling to work flows in Leeds from 2011 data, from the National Propensity to Cycle Tool (NPCT)."}
grid.raster(readPNG("../figures/npct-leeds-screenshot.png"))
```



