---
title: "Accident hotspots in Leeds"
author: "Robin Lovelace"
date: "May 29, 2015"
output: pdf_document
---

This represents a preliminary analysis of accident hotspots
for cyclists in Leeds.
It is preliminary because it does not include 2014 data and because
it does not normalise by estimated exposure. Also, it does not
make the distinction between different types of cycling trips
(primarily commuter trips, by time of day). This has the advantage of
including more data points. The disadvantage is that it heavily reflects
where cycling is greatest in Leeds, in the north of the city.

All the analysis underlying this research is reproducible.
In addition, the data underlying the results can be accessed for free online
by following the code provided in the paper's repository: 
https://github.com/Robinlovelace/bikeR

In addition, key datasets have been placed on an interactive
web map so they can be explored.

The 


```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
